---
title: "Linear_regression_model_ML_example"
author: "Aayushi"
date: '2024-04-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R-markdown for a sample linear regression model to predict 

```{r packages}
library(dplyr)
library(tidyverse)
library(ggplot2)
```
#The csv contains the logcpm expression values of different genes belonging to different groups


```{r data}
#Loading in the COVID cases data 
covid_data <- read.csv("C:/Users/notay003/Downloads/r_data/owid-covid-data.csv")
View(covid_data)

#subsetting the dataframe to columns of interest and removing rows with NA

covid_data <- covid_data[,c(1,3,5,6,8,18,20,32,35,37,47,48,49)]
covid_cases_na_removed <- na.omit(covid_data)

```

```{r}
#Performing linear regression to test if the number of deaths are related to the hospital, icu admission and vacinated status of pateints 

#since linear regression is a subcategory of Supervised ML algorithm we would need to split our data into training and test dataset

#Usually the training to test ratio is 70:30 
#Random division of the samples into training and test datasets 

rows_covid <- nrow(covid_cases_na_removed)
train <- sample(rows_covid, 0.7*rows_covid)
test <- rows_covid %>% seq_len() %>% setdiff(train)
```


```{r, echo=T}
#our dataset had 2882 rows
cat("Total no of enteries:",rows_covid, "\nEnteries in train dataset:",length(train),"\nEnteries in test dataset:",length(test))
```
```{r}
#subsetting the dataset into train and test
traning_covid <- covid_cases_na_removed[train,]
test_covid <- covid_cases_na_removed[test,]
```


```{r}
#fitting a linear model
#the method used is the default 'qr' method which works by finding the best-fitting linear approximation to the data by minimizing the sum of the squares of the differences between the observed and predicted values

#here we are considering total_deaths as a dependent variable and 
linear_model = lm(data=traning_covid, total_deaths~icu_patients+hosp_patients+positive_rate+total_vaccinations)
```

#having a look at how well the model fits the data

```{r}
summary(linear_model)
```

#the model appears to be good, as indicated by the high R-squared value (0.9551) and the statistically significant coefficients

```{r}
predicted_deaths = predict(linear_model,test_covid)

#extracting the total deaths from the original dataframe and plotting it against the predicted deaths obtained from our model 


results_covid = as.data.frame(predicted_deaths)
results_covid$Deaths_from_data = test_covid$total_deaths
results_covid$location <- test_covid$location
View(results_covid)

```

```{r,echo=T}
ggplot(results_covid,aes(x=predicted_deaths,y=Deaths_from_data,color=location))+geom_point()+
  theme_bw()+labs(title = "Linear Model for Covid Deaths",x="Predictde_Death",y="Known_Deaths")+
  theme(plot.title = element_text(size=10,face="bold"),
        axis.title = element_text(size=10))
```


```{r,echo=T}

#summerising the dataset by location for better plotting 
checking_for_sum <- results_covid %>% group_by(location) %>%summarise(Known_deaths=sum(Deaths_from_data),Predicted_deaths=sum(predicted_deaths))


ggplot(checking_for_sum,aes(x=Predicted_deaths,y=Known_deaths,color=location))+geom_point()+
  theme_bw()+labs(title = "Linear Model for Covid Deaths",x="Predictde_Death",y="Known_Deaths")+
  theme(plot.title = element_text(size=10,face="bold"),
        axis.title = element_text(size=10))+geom_abline(slope = 1, intercept = 0)#+guides(color='none')+geom_text_repel(aes(label=location),max.overlaps = Inf)

```

#As we can see that most of the points lie along the diagnol this means that our model prediction is very close to the original deaths from the stats



